Queue Aware Precoder Design for Space Frequency Resource Allocation

### 2014-07-20 Report/dual_approach.tex
Corrections after first reading

Let \me{\mbfa{\zeta}(b_k)} denotes the consensus entries corresponding to the \ac{BS} \me{b_k}. Let \me{\mbfa{\nu}^{\set{b_k}}} represents the stacked dual variables corresponding to the equality condition \me{\mbfa{\zeta}^{\set{b_k}} = \mbfa{\zeta}(b_k)} used in the subproblems. The equality constraint for the local and the consensus interference vector \me{\mbfa{\zeta}^{\set{b_k}} = \mbfa{\zeta}(b_k)} present in each subproblem is relaxed by the taking the partial Lagrangian for the constraint as
\begin{IEEEeqnarray}{rCl} \label{eqn-dual-2} \label{eqn-dual-1}
\underset{\substack{\gamma_{l,k,n}, \nu^{\set{b_k}}_{l,k,n,b} \\ \mvec{m}{l,k,n}, \beta_{l,k,n}, \zeta^{\set{b_k}}_{l,k,n,b}}}{\text{minimize}} &\quad & \| \tilde{\mbf{v}}_{b_k} \|_q + \mbfa{\nu}^{{\set{b_k}}T} \left ( \mbfa{\zeta}^{\set{b_k}} - \mbfa{\zeta}(b_k) \right ) \IEEEyessubnumber \label{eqn-dual-2a} \\
\text{subject to} && \beta_{l,k,n} \geq \sum_{\substack{j = 1\\j \neq l}}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{{b_k},k,n} \mvec{m}{j,k,n} |^2 \nonumber \\
&&\quad + \sum_{i \in \mc{U}_{b_k} \backslash \{k\}} \sum_{j = 1}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{{b_k},k,n} \mvec{m}{j,i,n} |^2 + \sum_{b \in \bar{\mc{B}}_{b_k}} \zeta^{\set{b_k}}_{l,k,n,b} \; + \; N_0\|\mvec{w}{l,k,n}\|^2 \IEEEyessubnumber \label{eqn-dual-1d} \\
&& \zeta^{\set{b_k}}_{\pr{l},\pr{k},n,{b_k}} \geq \sum_{k \in \mc{U}_b} \sum_{l = 1}^L |\mvec{w}{\pr{l},\pr{k},n}^\herm \mvec{H}{b_k,\pr{k},n} \mvec{m}{l,k,n} |^2, \; \forall \pr{k} \in \bar{\mc{U}}_{b_k}, \; \forall n \in \mc{C} \IEEEyessubnumber \label{eqn-dual-1e} \\
&\quad& \mbfa{\nu}^{\set{b_k}} \in \mathbb{R}^{n_{b_k}}_+, \; \eqref{eqn-8} \; \text{and} \; \eqref{eqn-primal-1c}. \IEEEyessubnumber \label{eqn-dual-1f}
\end{IEEEeqnarray}
The global interference variable \me{\mbfa{\zeta}(b_k)} in \eqref{eqn-dual-2a} can be dropped without affecting the optimal solution, since it is fixed for a given iteration. Instead of solving the dual decomposition problem defined in \eqref{eqn-dual-2}, we consider the \ac{ADMM} scheme which has better convergence properties as discussed in \cite{bertsekas1999nonlinear,boyd2011distributed}.

In order to limit the local interference assumptions \me{\zeta^{\set{b_k}}_{l,k,n,b}} at the \acp{BS} \me{b_k}, the \ac{ADMM} method augments a scaled quadratic penalty of the interference deviation between the local and consensus value for the interference from the \ac{BS} \me{b} as \me{\zeta_{l,k,n,b}} to the objective function. At optimality, the locally assumed and the consensus interference values will be equal,  providing no contribution to the objective function. The optimal step size use to update the dual variables is the scaling factor \me{\rho} used to scale the penalty term in the objective function \cite{boyd2011distributed}. Now, the subproblem at each \ac{BS} for the \me{\ith{i}} iteration is give by
\begin{IEEEeqnarray}{rCl} \label{eqn-dual-3}
\underset{\substack{\gamma_{l,k,n}, \mvec{m}{l,k,n}, \\ \beta_{l,k,n}, \zeta^{\set{b_k}(i)}_{l,k,n,b}}}{\text{minimize}} &\quad & \| \tilde{\mbf{v}}_{b_k} \|_q + \mbfa{\nu}^{{\set{b_k}(i-1)}T} \mbfa{\zeta}^{\set{b_k}(i)} + \frac{\rho}{2} \Big \| \, \underbrace{\mbfa{\zeta}^{(i-1)}(b_k)}_{\text{consensus values}} - \underbrace{\mbfa{\zeta}^{\set{b_k}(i)}}_{\text{local values}} \, \Big \|^2_2 \IEEEyessubnumber \label{eqn-dual-3a} \\
\text{subject to} & \quad & \eqref{eqn-dual-1d} - \eqref{eqn-dual-1f} \IEEEyessubnumber \label{eqn-dual-3b},
\end{IEEEeqnarray}
where the superscript \me{i} represents the current iteration or the information exchange index and \me{\mbfa{\zeta}^{(i-1)}} denotes the updated global interference level from the \me{\ith{(i-1)}} information exchange of the local interference vector \me{\mbfa{\zeta}^{\set{b}(i-1)}, \forall b \in \mc{B}}.

>>>>

Let \me{\mbfa{\zeta}(b_k)} denotes the consensus entries corresponding to the \ac{BS} \me{b_k}. Let \me{\mbfa{\nu}^{\set{b_k}}} represents the stacked dual variables corresponding to the equality condition \me{\mbfa{\zeta}^{\set{b_k}} = \mbfa{\zeta}(b_k)} used in the subproblems. The equality constraint for the local and the consensus interference vector \me{\mbfa{\zeta}^{\set{b_k}} = \mbfa{\zeta}(b_k)} present in each subproblem is relaxed by the taking the partial Lagrangian for the constraint as
\begin{IEEEeqnarray}{rCl} \label{eqn-dual-2} \label{eqn-dual-1}
\underset{\substack{\gamma_{l,k,n}, \nu^{\set{b_k}}_{l,k,n,b},\mvec{W}{k,n} \\ \mvec{M}{k,n}, \beta_{l,k,n}, \zeta^{\set{b_k}}_{l,k,n,b}}}{\text{minimize}} &\quad & \| \tilde{\mbf{v}}_{b_k} \|_q + \mbfa{\nu}^{{\set{b_k}}T} \left ( \mbfa{\zeta}^{\set{b_k}} - \mbfa{\zeta}(b_k) \right ) \IEEEyessubnumber \label{eqn-dual-2a} \\
\text{subject to} && \beta_{l,k,n} \geq \sum_{\substack{j = 1\\j \neq l}}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{{b_k},k,n} \mvec{m}{j,k,n} |^2 \nonumber \\
&&\quad + \sum_{i \in \mc{U}_{b_k} \backslash \{k\}} \sum_{j = 1}^L |\mvec{w}{l,k,n}^\herm \mvec{H}{{b_k},k,n} \mvec{m}{j,i,n} |^2 + \sum_{b \in \bar{\mc{B}}_{b_k}} \zeta^{\set{b_k}}_{l,k,n,b} \; + \; N_0\|\mvec{w}{l,k,n}\|^2 \IEEEyessubnumber \label{eqn-dual-1d} \\
&& \zeta^{\set{b_k}}_{\pr{l},\pr{k},n,{b_k}} \geq \sum_{k \in \mc{U}_b} \sum_{l = 1}^L |\mvec{w}{\pr{l},\pr{k},n}^\herm \mvec{H}{b_k,\pr{k},n} \mvec{m}{l,k,n} |^2, \; \forall \pr{k} \in \bar{\mc{U}}_{b_k}, \; \forall n \in \mc{C} \IEEEyessubnumber \label{eqn-dual-1e} \\
&\quad& \mbfa{\nu}^{\set{b_k}} \in \mathbb{R}^{n_{b_k}}_+, \; \eqref{eqn-8} \; \text{and} \; \eqref{eqn-primal-1c}. \IEEEyessubnumber \label{eqn-dual-1f}
\end{IEEEeqnarray}
The global interference variable \me{\mbfa{\zeta}(b_k)} in \eqref{eqn-dual-2a} can be dropped without affecting the optimal solution, since it is fixed for a given iteration. Instead of solving the dual decomposition problem defined in \eqref{eqn-dual-2}, we consider the \ac{ADMM} scheme which has better convergence properties as discussed in \cite{bertsekas1999nonlinear,boyd2011distributed}.

In order to limit the local interference assumptions \me{\zeta^{\set{b_k}}_{l,k,n,b}} at the \acp{BS} \me{b_k}, the \ac{ADMM} method augments a scaled quadratic penalty of the interference deviation between the local and consensus value for the interference from the \ac{BS} \me{b} as \me{\zeta_{l,k,n,b}} in the objective function. At optimality, the locally assumed and the consensus interference values will be equal,  providing no contribution to the objective function. The optimal step size used to update the dual variables is the scaling factor \me{\rho} used to scale the penalty term in the objective function \cite{boyd2011distributed}. Now, the subproblem at each \ac{BS} for the \me{\ith{i}} iteration is given by
\begin{IEEEeqnarray}{rCl} \label{eqn-dual-3}
\underset{\substack{\gamma_{l,k,n}, \mvec{M}{k,n},\mvec{W}{k,n} \\ \beta_{l,k,n}, \zeta^{\set{b_k}(i)}_{l,k,n,b}}}{\text{minimize}} &\quad & \| \tilde{\mbf{v}}_{b_k} \|_q + \mbfa{\nu}^{{\set{b_k}(i-1)}T} \mbfa{\zeta}^{\set{b_k}(i)} + \frac{\rho}{2} \Big \| \, \underbrace{\mbfa{\zeta}^{(i-1)}(b_k)}_{\text{consensus values}} - \underbrace{\mbfa{\zeta}^{\set{b_k}(i)}}_{\text{local values}} \, \Big \|^2_2 \IEEEyessubnumber \label{eqn-dual-3a} \\
\text{subject to} & \quad & \eqref{eqn-dual-1d} - \eqref{eqn-dual-1f} \IEEEyessubnumber \label{eqn-dual-3b},
\end{IEEEeqnarray}
where the superscript \me{i} represents the current iteration or the information exchange index and \me{\mbfa{\zeta}^{(i-1)}} denotes the updated global interference level from the \me{\ith{(i-1)}} information exchange of the local interference vector \me{\mbfa{\zeta}^{\set{b}(i-1)}, \forall b \in \mc{B}}.

### 2014-07-20 Report/introduction.tex
Corrections after first reading

The last mile wireless connectivity poses significant bottleneck in the overall data traffic in the interconnected networks. The main challenges in the wireless networks are due to the scarcity of the available resources either in terms of power or spectrum usage and the complexity of the receiver algorithms which is proportional to the battery drain. In order to over come the receiver complexity, \ac{OFDM} based transmissions are introduced for the wideband transmissions. To improve the data rate, multiple antennas are installed at the \ac{BS} and at the user terminals to avail the additional freedom in the form of spatial dimension. The inclusion of \ac{MIMO} technique in the wireless network provides increased data rate or lower outage for the same transmission power and the channel use.

In a network with multiple \acp{BS} serving multiple users (\acs{MU}), the main driving factor for the transmission are the packets waiting at each \ac{BS} corresponding to the different users present in the network. These available packets are transmitted over the shared wireless resources in an efficient manner with the various system constraints. In this work, we consider the user allocation over the space-frequency resources provided by the \ac{MU-MIMO} \ac{OFDM} framework in the downlink broadcast transmission to reduce the number of queued packets corresponding to the users at the \acp{BS}. Since the space-frequency resources are shared by multiple users associated with different \acp{BS}, the problem can be considered as a resource allocation to minimize the number of backlogged packets at the \acp{BS}.

In general, the resource allocation problems are solved by assigning a binary variable to each user indicating the presence or the absence on a particular resource. In contrast to that, we use the transmit beamformers, which are the complex vectors, as a decision variable in determining the presence or the absence of a user on a particular resource. The purpose of using the transmit beamformers for the scheduling is two fold. Firstly, it determines the transmission rate on a certain resource and secondly, by making the transmit beamformer to be a zero vector, the corresponding user will not be scheduled on a certain resource. The queue minimizing network optimization objective is used to design beamformers across the coordinating \acp{BS}, since the transmissions are guided by the available backlogged packets. To achieve the best performance, we propose a joint resource allocation scheme over the space and frequency dimensions among the coordinating \acp{BS} to minimize the time that the packets stay in queues prior to transmission, and, hence, to avoid packet drops as an indirect objective.

The queue minimizing precoder designs are nothing but the \ac{WSRM} problem with the additional rate constraint determined by the number of backlogged packets for each user in the system. The topic of \ac{MIMO} \ac{BC} precoder design has been studied extensively with different performance criteria in the literature. Due to the nonconvex nature of the linear \ac{MIMO} \ac{BC} precoder design problem, it is solved by the sequence of approximated convex problems. In \cite{sin_algorithm}, the \ac{MIMO} rate expression is linearized around a fixed point in order to solve the \ac{DC} problem as a concave maximization problem in an iterative manner. Similar approach of solving the \ac{WSRM} problem via arithmetic-geometric inequality is proposed in \cite{tran2012fast}, providing better convergence rate compared to the former.

The relation between the achievable capacity and the \ac{MSE} for the received symbol by using the fixed \ac{MMSE} receivers as shown in \cite{viswanath1999optimal,mse_duality} are also used to solve the \ac{WSRM} objective. In \cite{christensen2008weighted,wmmse_shi}, the \ac{WSRM} problem is reformulated via \ac{MMSE}, casting the problem as a convex one for fixed \ac{MMSE} receivers. In this way, the original problem is expressed in terms of the \ac{MSE} weight, precoders, and decoders. Then the problem is solved using an alternating optimization method, i.e., finding a set of variables while the remaining others are fixed. Additional rate constraints based on the \ac{QoS} requirements are included in the \ac{WSRM} objective are solved via \ac{MSE} reformulation in \cite{kaleva2013primal}.

@@

The outline of this paper is as follows. In Section \ref{sec-2-3.2}, we propose the system model and the problem formulation for the current queue minimizing precoder design. Existing and the proposed precoder designs for the queue minimizing objective is discussed elaborately in Section \ref{sec-3}. The distributed solutions are outlined in Section \ref{sec-4} followed by the simulation results for both centralized and the distributed algorithms are provided in Section \ref{sec-5}. Conclusions are drawn in Section \ref{sec-6}.

>>>>

The last mile wireless connectivity poses significant bottleneck in the overall data traffic for the interconnected networks. The main challenges in the wireless networks are due to the scarcity of the available resources either in terms of power or spectrum usage and the complexity of the receiver algorithms which is proportional to the mobile battery drain. In order to over come the receiver complexity, \ac{OFDM} based transmissions are introduced for the wideband transmissions. To improve the data rate, multiple antennas are installed at the \ac{BS} and at the user terminals to avail the additional freedom in the form of spatial dimension. The inclusion of \ac{MIMO} technique in the wireless network provides increased data rate or lower outage for the same transmission power and the channel use.

In a network with multiple \acp{BS} serving multiple users (\acs{MU}), the main driving factor for the transmission are the packets waiting at each \ac{BS} corresponding to the different users present in the network. These available packets are transmitted over the shared wireless resources in an efficient manner by including the system limitations and constraints. In this work, we consider the user allocation over the space-frequency resources provided by the \ac{MU-MIMO} \ac{OFDM} framework in the downlink broadcast transmission to reduce the number of queued packets corresponding to the users at the \acp{BS}. Since the space-frequency resources are shared by multiple users associated with different \acp{BS}, the problem can be considered as a resource allocation to minimize the number of backlogged packets at the \acp{BS}.

In general, the resource allocation problems are solved by assigning a binary variable to each user indicating the presence or the absence on a particular resource. In contrast to that, we use the transmit beamformers, which are the complex vectors, as a decision variable in determining the presence or the absence of a user on a particular resource. The purpose of using the transmit beamformers for the scheduling is two fold. Firstly, it determines the transmission rate on a certain resource and secondly, by making the transmit beamformer to be a zero vector, the corresponding user will not be scheduled on a certain resource. The queue minimizing network optimization objective is used to design the beamformers across the coordinating \acp{BS}, since the transmissions are guided by the available backlogged packets. To achieve the best performance, we propose a joint resource allocation scheme over the space and frequency dimensions among the coordinating \acp{BS} to minimize the time that the packets stay in the queues prior to the transmission, and, hence, to avoid packet drops as an indirect objective.

The queue minimizing precoder designs are nothing but the \ac{WSRM} problem with the additional rate constraint determined by the number of backlogged packets for each user in the system. The topic of \ac{MIMO} \ac{BC} precoder design has been studied extensively with different performance criteria in the literature. Due to the nonconvex nature of the linear \ac{MIMO} \ac{BC} precoder design problem, it is usually solved by the sequence of approximated convex problems. In \cite{sin_algorithm}, the \ac{MIMO} rate expression is linearized around a fixed point in order to solve the \ac{DC} problem as a concave maximization problem in an iterative manner. Similar approach of solving the \ac{WSRM} problem via arithmetic-geometric inequality is proposed in \cite{tran2012fast}, providing better convergence rate compared to the former.

The relation between the achievable capacity and the \ac{MSE} for the received symbol by using the fixed \ac{MMSE} receivers as shown in \cite{viswanath1999optimal,mse_duality} are also used to solve the \ac{WSRM} objective. In \cite{christensen2008weighted,wmmse_shi}, the \ac{WSRM} problem is reformulated via \ac{MSE}, casting the problem as a convex one for fixed \ac{MMSE} receivers. In this way, the original problem is expressed in terms of the \ac{MSE} weight, precoders, and decoders. Then the problem is solved using an alternating optimization method, i.e., finding a set of variables while the remaining others are fixed. Additional rate constraints based on the \ac{QoS} requirements are included in the \ac{WSRM} objective are solved via \ac{MSE} reformulation in \cite{kaleva2013primal}.

@@

The outline of this paper is as follows. In Section \ref{sec-2-3.2}, we propose the system model and the problem formulation for the current queue minimizing precoder design. Existing and the proposed precoder designs for the queue minimizing objective are discussed elaborately in Section \ref{sec-3}. The distributed solutions are outlined in Section \ref{sec-4} followed by the simulation results for both centralized and the distributed algorithms are provided in Section \ref{sec-5}. Conclusions are drawn in Section \ref{sec-6}.

### 2014-07-20 Report/joint_wsrm_formulation.tex
Corrections after first reading

The queue minimizing algorithms are discussed extensively in the networking literature for providing congestion-free routing between any two nodes\footnote{routers or user terminals} in the network. It is achieved by finding a control policy considering the differential backlogs of the packets present in the source and the destination node, where control policy refers to the resource or rate allocation strategy. The control policy satisfying the condition is known as \emph{backpreassure algorithm}, discussed extensively in \cite{tassiulas,georgiadis2006resource,neely2010stochastic}. The algorithm assumes the physical layer rate as a fixed entity due to the wired infrastructure. In order to use the algorithm in the wireless network, the user rate variable \me{t_{k}} should be modeled in accordance with the wireless backbone network.

The \ac{Q-WSRM} scheme is formulated by extending the \emph{backpreassure algorithm} for the \ac{MIMO} \ac{OFDM} network, where multiple \acp{BS} act like a source nodes and the user terminals as a receiver nodes. The user rate variables \me{t_k} are obtained by the design of the transmit and the receive precoders to maximize the weighted sum rate. The weights used in the precoder design algorithm should be a function of the number of backlogged packets corresponding to each user. In order to find the weights, we adopt the Lyapunov drift minimization as discussed in \cite{neely2010stochastic}, where the Lyapunov function is given by the squared sum of the number backlogged packets as
\begin{equation}
\mathrm{L}\sset{\mbf{Q}(i)} = \frac{1}{2} \, \sum_{k \in \mc{U}} Q_k(i),
\end{equation}
where \me{\mbf{Q}(i)} denotes the stacked user queues at the \me{\ith{i}} slot. The Lyapunov function provides a measure of congestion in the system, as discussed in \cite[Ch. 3]{neely2010stochastic}. Now the Lyapunov function drift is given by
\begin{IEEEeqnarray}{rCl}\label{eqn-3.1}
\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} &=& \frac{1}{2} \left [ \sum_{k \in \mc{U}} \, \Big ( \left [ Q_k(i) - t_k(i) \right ]^+ + \lambda_k(i) \Big )^2 - Q^2_k(i) \right ] \IEEEyessubnumber \label{eqn-3.1.0} \\
&\leq& \sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} + \sum_{k \in \mc{U}} Q_k(i) \set{ \lambda_k(i) - t_k(i) }, \IEEEyessubnumber \label{drift-exp}
\end{IEEEeqnarray}
where the inequality is due to the upper bound
\begin{equation}
\sset{\max(Q-t,0) + \lambda)}^2 \leq Q^2 + t^2 + \lambda^2 + 2 Q (\lambda - t).
\end{equation}
In order to minimize the number of backlogged packets at each instant, minimization of the Lyapunov drift \eqref{eqn-3.1} is carried over all possible control decisions in the form of transmission rates \me{t_k} to users in the system. The Lyapunov drift conditioned on the current backlogged packets \me{\mbfa{Q}(i)} is given by
\begin{IEEEeqnarray}{rCl}
\underset{\mbf{t}}{\text{minimize}} && \Delta(\mbf{Q}(i)) \triangleq \mathbb{E} \set{\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} \vert \mbf{Q}(i)} \IEEEyessubnumber \\
 &\leq& \mathbb{E} \set {\sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} \vert \mbf{Q}(i)} + \sum_{k \in \mc{U}} Q_k(i) A_k(i) + \mathbb{E}\set{\sum_{k \in \mc{U}} Q_k(i) t_k(i)  \vert \mbf{Q}(i)}, \IEEEyessubnumber
\end{IEEEeqnarray}
where the second term on the r.h.s is due to i.i.d assumption on the arrivals. We can bound the first term on the r.h.s by a constant \me{B} for all possible control actions taken for a given channel condition \cite{neely2010stochastic}, the resulting problem can be given as
\begin{equation}
\underset{t_k}{\text{maximize}} \quad \sum_{k \in \mc{U}} \, Q_k(i) \, t_k(i),
\label{backpreassure}
\end{equation}
where the expectation operator is dropped since it is maximized by maximizing the function inside in an opportunistic manner.

>>>>

The queue minimizing algorithms are discussed extensively in the networking literature for providing a congestion-free routing between any two nodes\footnote{routers or user terminals} in the network. One such algorithm used extensively is the \emph{backpreassure algorithm}, discussed in detail in \cite{tassiulas,georgiadis2006resource,neely2010stochastic}. The algorithm determines an optimal control policy in the form of rate or resource allocation for the nodes in the network by considering the differential backlogged packets between the source and the destination nodes.
Even though the algorithm is primarily designed for the wired infrastructure, it can be extended to the wireless networks by designing the user rate variable \me{t_k} in accordance to the wireless network.

The \ac{Q-WSRM} formulation extends the \emph{backpreassure algorithm} to the \ac{MIMO} \ac{OFDM} framework, in which the multiple \acp{BS} acts as the source nodes and the user terminals as the receiver nodes. The control policy in the form of transmit precoders are designed with the objective of minimizing the number of queued packets waiting at the \acp{BS}. The weights used in the precoder design algorithm should be a function of the number of backlogged packets corresponding to each user. In order to find the weights, we adopt the Lyapunov drift minimization as discussed in \cite{neely2010stochastic}, where the Lyapunov function is given by the squared sum of the number backlogged packets as
\begin{equation}
\mathrm{L}\sset{\mbf{Q}(i)} = \frac{1}{2} \, \sum_{k \in \mc{U}} Q_k(i),
\end{equation}
where \me{\mbf{Q}(i)} denotes the stacked user queues at the \me{\ith{i}} slot. The Lyapunov function provides a measure of congestion in the system, as discussed in \cite[Ch. 3]{neely2010stochastic}. Now the Lyapunov function drift is given by
\begin{IEEEeqnarray}{rCl}\label{eqn-3.1}
\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} &=& \frac{1}{2} \left [ \sum_{k \in \mc{U}} \, \Big ( \left [ Q_k(i) - t_k(i) \right ]^+ + \lambda_k(i) \Big )^2 - Q^2_k(i) \right ] \IEEEyessubnumber \label{eqn-3.1.0} \\
&\leq& \sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} + \sum_{k \in \mc{U}} Q_k(i) \set{ \lambda_k(i) - t_k(i) }, \IEEEyessubnumber \label{drift-exp}
\end{IEEEeqnarray}
where the inequality is due to the upper bound
\begin{equation}
\sset{\max(Q-t,0) + \lambda)}^2 \leq Q^2 + t^2 + \lambda^2 + 2 Q (\lambda - t).
\end{equation}
In order to minimize the number of backlogged packets at each instant, minimization of the Lyapunov drift \eqref{eqn-3.1} is carried over all possible control decisions in the form of transmission rates \me{t_k} to users in the system. The Lyapunov drift conditioned on the current backlogged packets \me{\mbfa{Q}(i)} is given by
\begin{IEEEeqnarray}{rCl}
\underset{\mbf{t}}{\text{minimize}} && \Delta(\mbf{Q}(i)) \triangleq \mathbb{E} \set{\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} \vert \mbf{Q}(i)} \IEEEyessubnumber \\
 &\leq& \mathbb{E} \set {\sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} \vert \mbf{Q}(i)} + \sum_{k \in \mc{U}} Q_k(i) A_k(i) + \mathbb{E}\set{\sum_{k \in \mc{U}} Q_k(i) t_k(i)  \vert \mbf{Q}(i)}, \IEEEyessubnumber
\end{IEEEeqnarray}
where the second term on the r.h.s is due to i.i.d assumption on the arrivals. We can bound the first term on the r.h.s by a constant \me{B} for all possible control actions taken for a given channel condition \cite{neely2010stochastic}, the resulting problem can be given as
\begin{equation}
\underset{t_k}{\text{maximize}} \quad \sum_{k \in \mc{U}} \, Q_k(i) \, t_k(i),
\label{backpreassure}
\end{equation}
where the expectation operator is dropped since it is maximized by maximizing the function inside in an opportunistic manner.
